### 2019-03-31 doc/draft-01-stokes/stokes-eig.tex

\documentclass[preprint,12pt]{article} 
\usepackage{b}
\usepackage[a4paper]{geometry}
\geometry{top=1.0in, bottom=1.0in, left=1.0in, right=1.0in}
\usepackage[small,it]{caption}
\usepackage{empheq}

\usepackage{subcaption}
\usepackage{amsmath}

\usepackage{cleveref}
\crefname{equation}{}{}
\Crefname{equation}{}{}
\crefname{eqnarray}{}{}
\Crefname{eqnarray}{}{}
\crefname{lemma}{lemma}{lemmas}
\crefname{lem}{lemma}{lemmas}
\crefname{cor}{corollary}{corollaries}
\crefname{thrm}{theorem}{theorems}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\bz{{\boldsymbol 0}}
\def\br{{\boldsymbol r}}
\def\bc{{\boldsymbol c}}
\def\bh{{\boldsymbol h}}
\def\bn{{\boldsymbol n}}
\def\bT{{\boldsymbol T}}
\def\bt{{\boldsymbol t}}
\def\xx{{\boldsymbol x}}
\def\bx{{\boldsymbol x}}
\def\yy{{\boldsymbol y}}
\def\by{{\boldsymbol y}}
\def\ss{{\boldsymbol s}}
\def\tt{{\boldsymbol t}}
\def\uu{{\boldsymbol u}}
\def\bu{{\boldsymbol u}}
\def\be{{\boldsymbol e}}
\def\bl{{\boldsymbol \ell}}
\def\ff{{\boldsymbol f}}
\def\bg{{\boldsymbol g}}
\def\GG{{\bf G}}
\def\Gbh{G^{\textrm{BH}}}
\def\Glap{G^{\textrm{L}}}
\def\Ghelm{G^{\textrm{H}}}
\def\Kfark{K^{\textrm{F}}}
\def\II{{\bf I}}
\def\TT{{\bf T}}
\def\bS{{\bf S}}
\def\bD{{\bf D}}
\def\bC{{\bf C}}
\def\bB{{\bf B}}
\def\cN{{\mathcal N}}
\def\cK{{\mathcal K}}
\def\cB{{\mathcal B}}
\def\cA{{\mathcal A}}
\def\cS{{\mathcal S}}
\def\cSk{{\mathcal S}_k}
\def\cD{{\mathcal D}}
\def\cDt{{\mathcal D}^{\intercal}}
\def\cDk{{\mathcal D}_{k}}
\def\cDkt{{\mathcal D}_{k}^{\intercal}}
\def\cW{{\mathcal W}}
\def\cJ{{\mathcal J}}
\def\cL{{\mathcal L}}
\def\cA{{\mathcal A}}
\def\cM{{\mathcal M}}
\def\bL{{\mathbb L}}
\def\bK{{\mathbf K}}
\def\sint{{\sin{(n \theta)}}}
\def\cost{{\cos{(n \theta)}}}
\def\jnkr{J_{n}(kr)}
\def\jnk{J_{n}(k)}
\def\jnpkr{J_{n}'(kr)}
\def\jnppkr{J_{n}''(kr)}
\def\jnppk{J_{n}''(k)}
\def\ddr{\frac{d}{dr}}

\def\bsigma{{\boldsymbol \sigma}}
\def\bpsi{{\boldsymbol \psi}}
\def\bmu{{\boldsymbol \mu}}
\def\bnu{{\boldsymbol \nu}}
\def\brho{{\boldsymbol \rho}}
\def\btau{{\boldsymbol \tau}}
\def\bigo{\mathcal{O}}
\def\littleo{o}

\def\Im{{\textrm{\textup{Im}}}}

\def\epsflam{{\epsilon_{\texttt{FLAM}}}}
\def\epscheb{{\epsilon_{\texttt{CHEB}}}}

\newcommand\dwdn{\frac{\partial w}{\partial n}}
\newcommand{\cI}{\mathcal I}
\newcommand{\cC}{\mathcal C}
\newcommand{\bbmat}{\begin{bmatrix}}
\newcommand{\ebmat}{\end{bmatrix}}

\newcommand{\note}[1]{ {\color{red} #1 }}

\def\pv{\textrm{P.V.} }

\newtheorem{remark}{Remark}
\newtheorem{proposition}{Proposition}
\newtheorem{lem}{Lemma}
\newtheorem{cor}{Corollary}
\newtheorem{thrm}{Theorem}
\newtheorem{definition}{Definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{A boundary integral equation approach to
  computing eigenvalues of the Stokes operator}
%A Fredholm integral equation approach to
%  computing Stokes eigenvalues in the plane
\author{Travis Askham and Manas Rachh}
\date{\today}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abstract{The boundary integral equation (BIE) framework
  provides robust and scalable eigenvalue computations
  due to (a) the reduction in the dimension of the problem to
  be discretized and (b)
  the absence of high frequency ``pollution'' when using
  a Green's function to represent propagating waves.
  %
  In this paper, we detail the theoretical justification
  for a BIE approach to the Stokes eigenvalue problem on
  simply and multiply-connected planar domains, which entails
  a treatment of the uniqueness theory for oscillatory
  Stokes equations on exterior domains. 
  %
  Then, using well-established techniques for
  discretizing BIEs,
  we present numerical results which confirm
  the analytical claims of the paper and demonstrate
  the efficiency of the overall approach.
}

\input{01intro}
\input{02prelim}
\input{03analysis}
\input{04numerical}
\input{05conclusion}

\appendix

\input{appendix}
\bibliographystyle{plain}
\bibliography{refs}

\end{document}